---
title: "ESM 206 Lab 5"
author: "Claudia Flores"
date: "10/28/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## Objectives:

- Getting counts for different groups
- Use {lubridate} to parse dates
- Find confidence intervals and do t-tests with the t.test()
- Heatmap with geom_tile()

```{r, include = FALSE}
# Attach packages

library(tidyverse)
library(here)
library(janitor)

```

```{r}
# Import lobster data

lobster_abundance <- read_csv(here::here("data", "lobster_abundance.csv"), na = "-99999") %>% 
  clean_names()
```
Use the tidyr::uncount() function to expand a frequency variable
```{r}
lobster_tidy <- lobster_abundance %>%
  tidyr::uncount(lobster_count)
  
```

Yaya! Now each lobster has its own row. This is tidy format. 

Let's look at this a bunch of ways"
```{r}

ggplot(data = lobster_tidy, aes(x = site, y = size_mm)) +
  geom_jitter(aes(color = site), width = 0.2, alpha = 0.3)
```

```{r}

ggplot(data = lobster_tidy, aes(x = site, y = size_mm)) +
  geom_violin(aes(color = site))
```
```{r}
ggplot(data = lobster_tidy, aes(x = size_mm)) +
  geom_histogram(aes(fill = site)) +
  facet_wrap(~site, scales = "free")
```
```{r}
ggplot(data = lobster_tidy, aes(sample = size_mm)) +
  geom_qq(size = 0.5) +
  facet_wrap(~site, scales = "free")
```


Use 'lubridate' to parse dates and times:
Always use lubridate:: when using the lubridate package because there are similiar functions in other packages
Use class(df$column) to check character type or lapply(df,class)
I'm going to add a new column that contains date as an actual date 
```{r}
lobster_date <- lobster_tidy %>% 
  mutate(date_new = lubridate::mdy(date))

```

Parse 'Date' column to get the different pieces (month,year) separated
```{r}
lobster_parse_date <- lobster_date %>%
  mutate(obs_month = lubridate::month(date_new, label = TRUE),
  obs_year = lubridate::year(date_new))

#Label knows that there is an order to them, so will put the first month in the beginning 
```

Count lobsters by different groupings...

Let's say I want to count the number of lobsters by year and month
```{r}
lobster_ym <- lobster_parse_date %>% 
  count(obs_year, obs_month)

lobster_ym

lobster_ysite <- lobster_parse_date %>%
  count(obs_year, site)

lobster_ysite

lobster_site <- lobster_parse_date %>%
  count(site)

lobster_site

#group_by + summarise + n()

lobster_summary <- lobster_parse_date %>% 
  group_by(site) %>% 
  summarize(mean_size = mean(size_mm, na.rm = TRUE),
  sd_size = sd(size_mm, na.rm = TRUE),
  sample_n = n())

# Note: tally is like count, you just have to do group_by before it

```

Confidence Intervals and t-tests

Use the t-test function to find the confidence intervals and perform t-tests
filter is subsetting by rows
```{r}
ivee_lobsters <- lobster_tidy %>% 
  filter(site == "IVEE") %>% 
  pull(size_mm)

# pull() will create a vector

t.test(ivee_lobsters)
```


Two sample t-test: Is there a significant difference in mean lobster lengths between Naples Reef and Mohwak Reef?
```{r}

napl_sample <- lobster_tidy %>%
  filter(site == "NAPL") %>% 
  pull(size_mm)


mohk_sample <- lobster_tidy %>%
  filter(site == "MOHK") %>% 
  pull(size_mm)


mohk_napl_ttest <- t.test(napl_sample, mohk_sample)

mohk_napl_ttest
```

Mean lobster size differ significantly between Mohawk and Naples reefs
(t{'r mohk_napl_ttest$parameter'} = 19.849)...

Maybe an easier way...
```{r}
lobster_2sample <- lobster_tidy %>%
  filter(site %in% c("NAPL","MOHK"))

ttest_2 <- t.test(size_mm ~ site, data = lobster_2sample)
ttest_2
```

Make a geom_tile heatmap
```{r}
ggplot(data = lobster_ysite, aes(x = obs_year, y = site)) +
  geom_tile(aes(fill = n))
```

